\section{Conclusion}

We presented the design, implementation, and systematic benchmarking of
\textsc{ChainedBatchStateTransition}, a Circom circuit for proving
batched Sparse Merkle Tree state transitions with Poseidon hashing
under Groth16.

Our principal contributions are:

\begin{enumerate}
\item An exact constraint scaling law, $C_{\text{tx}} = 1{,}038 \cdot (d+1)$,
  verified with zero residual error across seven configurations spanning
  three tree depths and three batch sizes.

\item Empirical evidence that the production target of depth~32,
  batch~64 ($\sim$2.19M constraints) is provable in 14--35~seconds
  with Rapidsnark, well within a 60-second enterprise latency budget.

\item Quantitative reconciliation with production systems: our
  per-transaction cost of 34,254~constraints at depth~32 is consistent
  with Hermez ($0.59\times$, excluding signature and token logic) and
  Tornado Cash ($3.1\text{--}4.4\times$ a single Merkle proof, consistent
  with dual-path verification).
\end{enumerate}

The results reject the na\"ive 100,000-constraint target for batch-64
circuits but confirm the feasibility of enterprise validium batch
processing when paired with native Groth16 provers. The derived
constraint formula enables exact capacity planning: given a prover's
throughput and a latency budget, the maximum batch size can be computed
directly.

\subsection{Future Work}

Three extensions are planned within the Basis Network research pipeline:

\begin{itemize}
\item \textbf{Adversarial testing} (Stage~3): fault injection,
  malicious witness inputs, and constraint under-specification attacks
  to verify circuit soundness under adversarial conditions.

\item \textbf{Rapidsnark benchmarking}: direct measurement of native
  prover performance across the same configurations to validate the
  extrapolated speedup factors.

\item \textbf{SMTProcessor comparison}: implementation and benchmarking
  of an equivalent circuit using circomlib's \texttt{SMTProcessor}
  template to quantify the constraint savings from shared intermediate
  computations.
\end{itemize}

The circuit design and constraint formula presented here feed directly
into the downstream formal verification pipeline (TLA+ specification
by the Logicist agent, implementation by the Architect agent, and
Coq certification by the Prover agent) of the Basis Network R\&D
laboratory.
