% abstract.tex -- Enterprise Node Orchestrator Paper (RU-V5)

\begin{abstract}
Zero-knowledge validium nodes for enterprise blockchain require orchestrating
a multi-phase pipeline---transaction ingestion, Sparse Merkle Tree state
updates, witness generation, ZK proof computation, data availability committee
attestation, and L1 settlement---while maintaining continuous availability and
zero data leakage. We present the design and empirical evaluation of a
pipelined, event-driven state machine orchestrator for the Basis Network
enterprise ZK validium node on Avalanche. The orchestrator implements six
states with 17 valid transitions, enabling concurrent transaction ingestion
during proof generation and L1 submission through three decoupled processing
loops. Benchmarks across five configurations with 30--50 replications each
demonstrate that orchestration overhead (batch formation plus witness
generation) consumes 593~ms per 8-transaction batch---just 0.66\% of the
90-second end-to-end budget. ZK proof generation dominates latency at
64--85\% of total cycle time across all tested backends. Pipelined execution
achieves a 1.29$\times$ throughput improvement over sequential processing for
64-transaction batches with the rapidsnark backend, yielding 4.37~tx/s
sustained throughput. The results identify a critical constraint: the snarkjs
WASM prover exceeds the 90-second target at batch size 64 (156.9~s per batch),
requiring either a native prover (rapidsnark: 18.9~s) or batch size reduction
to 16 (31.4~s) for production deployment. All findings are parameterized for
downstream TLA+ formalization and implementation.
\end{abstract}
