% results.tex -- Sparse Merkle Trees with Poseidon Hash

\section{Results}
\label{sec:results}

All experiments were conducted on a Windows~11 x64 workstation running Node.js v22.13.1 with circomlibjs v0.1.7. Results are reported from the BN128 scalar field configuration.

\subsection{Hash Function Performance}

Table~\ref{tab:hash-performance} presents the native JavaScript execution performance of Poseidon and MiMC over the BN128 scalar field, measured over 1{,}000 repetitions after 100 warmup iterations.

\begin{table}[htbp]
\centering
\caption{Hash function performance in native JavaScript (BN128 field).}
\label{tab:hash-performance}
\begin{tabular}{lrr}
\hline
\textbf{Configuration} & \textbf{Latency ($\mu$s/hash)} & \textbf{Throughput (hash/s)} \\
\hline
Poseidon 2-to-1 & 56.09 & 17{,}830 \\
Poseidon 1-input & 40.44 & 24{,}728 \\
MiMC-Feistel ($x^7$, 91 rounds) & 278.55 & 3{,}590 \\
Poseidon chain-32 & 53.54/hash & 18{,}677 \\
\hline
\end{tabular}
\end{table}

Poseidon achieves a \textbf{4.97$\times$ speedup} over MiMC in native JavaScript execution. This is consistent with the theoretical ratio of constraint counts ($340/240 = 1.42\times$) amplified by MiMC's higher round count (91 rounds versus Poseidon's ${\sim}65$ effective rounds) and the quadratic cost of BigInt modular exponentiation for $x^7$ versus $x^5$.

A full depth-32 Merkle path verification requires 32 sequential Poseidon hashes, measured at 1.71\,ms per chain ($53.54\,\mu$s per individual hash within the chain). The slight reduction in per-hash latency within chains compared to isolated hashes (53.54 vs.\ 56.09\,$\mu$s) is attributable to JIT optimization of the hot loop.

\subsection{SMT Insert Latency}

Table~\ref{tab:insert-latency} presents insert latency across four tree sizes, demonstrating the O($d$) complexity where $d = 32$ is the tree depth.

\begin{table}[htbp]
\centering
\caption{SMT insert latency by tree size (50 samples per configuration).}
\label{tab:insert-latency}
\begin{tabular}{rrrrr}
\hline
\textbf{Entries} & \textbf{Mean (ms)} & \textbf{P50 (ms)} & \textbf{P95 (ms)} & \textbf{Stddev (ms)} \\
\hline
100 & 1.877 & 1.864 & 2.087 & 0.134 \\
1{,}000 & 1.788 & 1.780 & 1.997 & 0.095 \\
10{,}000 & 1.792 & 1.777 & 1.946 & 0.085 \\
100{,}000 & 1.825 & 1.811 & 2.014 & 0.098 \\
\hline
\end{tabular}
\end{table}

Insert latency is effectively constant across tree sizes at approximately 1.8\,ms, confirming the O($d$) complexity prediction. The latency is dominated by 32 Poseidon hash computations (${\sim}1.7$\,ms) plus hash map lookups for sibling retrieval. The slight variation across sizes (1.788--1.877\,ms) falls within one standard deviation and is attributable to JIT warmup effects and V8 garbage collection.

\subsection{Proof Generation Time}

Table~\ref{tab:proof-gen} shows proof generation times, which involve tree traversal without hash computation.

\begin{table}[htbp]
\centering
\caption{SMT proof generation time by tree size (50 samples per configuration).}
\label{tab:proof-gen}
\begin{tabular}{rrr}
\hline
\textbf{Entries} & \textbf{Mean (ms)} & \textbf{P95 (ms)} \\
\hline
100 & 0.012 & 0.029 \\
1{,}000 & 0.014 & 0.016 \\
10{,}000 & 0.014 & 0.016 \\
100{,}000 & 0.018 & 0.021 \\
\hline
\end{tabular}
\end{table}

Proof generation is extremely fast ($< 0.02$\,ms at 100K entries) because it only traverses the tree to collect the 32 sibling hashes---no hash computation is required. This is 278$\times$ faster than the 5\,ms target. The marginal increase from 0.012\,ms (100 entries) to 0.018\,ms (100K entries) reflects the larger hash map size increasing lookup times.

\subsection{Proof Verification Time}

Table~\ref{tab:proof-verify} presents verification times, which require 32 sequential Poseidon hashes.

\begin{table}[htbp]
\centering
\caption{SMT proof verification time by tree size (50 samples per configuration).}
\label{tab:proof-verify}
\begin{tabular}{rrr}
\hline
\textbf{Entries} & \textbf{Mean (ms)} & \textbf{P95 (ms)} \\
\hline
100 & 1.685 & 1.788 \\
1{,}000 & 1.677 & 1.758 \\
10{,}000 & 1.719 & 1.905 \\
100{,}000 & 1.744 & 1.869 \\
\hline
\end{tabular}
\end{table}

Verification time is constant at approximately 1.7\,ms across all tree sizes, as expected (always 32 Poseidon hashes regardless of population). This is consistent with the chain-32 benchmark (1.71\,ms) from Section~\ref{sec:results}.A. The P95 at 100K entries (1.869\,ms) is within the 2\,ms target with a 7\% margin.

\subsection{Memory Usage}

Table~\ref{tab:memory} summarizes memory consumption across tree sizes.

\begin{table}[htbp]
\centering
\caption{Memory usage by tree size.}
\label{tab:memory}
\begin{tabular}{rrrr}
\hline
\textbf{Entries} & \textbf{Nodes Stored} & \textbf{Heap (MB)} & \textbf{Bytes/Entry} \\
\hline
100 & 2{,}721 & 68.4 & $\sim$700{,}000 \\
1{,}000 & 23{,}867 & 112.8 & $\sim$115{,}000 \\
10{,}000 & 202{,}398 & 57.9 & $\sim$5{,}900 \\
100{,}000 & 1{,}712{,}112 & 233.9 & $\sim$2{,}400 \\
\hline
\end{tabular}
\end{table}

Node count scales at approximately 17$\times$ the entry count, consistent with each insert creating or updating up to 33 nodes (1 leaf + 32 internal path nodes). At 100K entries, memory usage of 233.9\,MB is well within the 2\,GB threshold. The amortized bytes-per-entry decreases with tree size due to fixed V8 runtime overhead being distributed across more entries. The anomalous 57.9\,MB reading at 10K entries is attributed to V8 garbage collection between benchmark runs (each tree size uses a fresh SMT instance).

\subsection{Benchmark Reconciliation}

Table~\ref{tab:reconciliation} compares our measurements against published reference values.

\begin{table}[htbp]
\centering
\caption{Benchmark reconciliation with published data.}
\label{tab:reconciliation}
\begin{tabular}{lllp{3.5cm}}
\hline
\textbf{Metric} & \textbf{Ours} & \textbf{Published} & \textbf{Assessment} \\
\hline
Poseidon vs MiMC & 4.97$\times$ & $\sim$2$\times$~\cite{semaphore2024v4} & Consistent (circuit vs.\ native) \\
SMT insert & 1.825\,ms & 1.92\,$\mu$s~\cite{reilabs2024largesmt} & Expected: JS vs.\ Rust \\
Proof generation & 0.018\,ms & 20--83\,$\mu$s~\cite{reilabs2024largesmt} & Comparable \\
Poseidon constraints & 240 & 240~\cite{guohanze2024benchmarking} & Exact match \\
\hline
\end{tabular}
\end{table}

The 950$\times$ gap between our JavaScript insert latency and Rust implementations is expected and not concerning: Rust uses native 256-bit arithmetic while JavaScript relies on BigInt with heap-allocated arbitrary-precision integers. The relevant comparison is against the stated performance targets, all of which pass.
