% conclusion.tex -- Batch Aggregation Paper (BNR-2026-004)

\section{Conclusion}

We presented a crash-safe batch aggregation protocol for enterprise ZK validium
nodes, combining a WAL-backed persistent queue with a HYBRID batch formation
strategy. Empirical evaluation across 17 configurations with 30 replications
each demonstrates 274,438~tx/min peak throughput, sub-millisecond batch
formation latency, 450/450 determinism, and 150/150 crash recovery success.

The primary contribution is the discovery and resolution of a silent transaction
loss bug through TLA+ formal verification. The bug---a premature WAL checkpoint
that creates a durability gap between batch formation and ZK proof
completion---was invisible to 150 crash recovery tests but found by TLC model
checking in under one second via a 6-step counterexample. The corrected protocol
(deferred checkpoint) is verified to satisfy all 6 safety invariants and 1
liveness property across 2,630 distinct states with zero errors.

This result demonstrates that formal verification and empirical testing are
complementary, not redundant, for crash-safe protocol design. We recommend that
ZK infrastructure projects adopt lightweight formal methods---TLA+ specifications
can be written in hours and model-checked in seconds, yet they catch bugs that
may never surface in testing until production deployment.

Future work includes adversarial testing (concurrent writers, WAL corruption
injection, large-batch stress), binary WAL format for high-throughput
deployments, and integration with the Sparse Merkle Tree and ZK proving
pipeline for end-to-end batch-to-proof benchmarks.
