% abstract.tex -- Sparse Merkle Trees with Poseidon Hash

\begin{abstract}
Enterprise blockchain systems face a fundamental tension between verifiability and data privacy. Zero-knowledge validium architectures resolve this by maintaining state off-chain and submitting validity proofs to a Layer~1 chain, but their performance depends critically on the state commitment structure. This paper presents the design, implementation, and empirical evaluation of a depth-32 Sparse Merkle Tree (SMT) using the Poseidon hash function over the BN128 scalar field, purpose-built for the Basis Network enterprise ZK validium node on Avalanche. We evaluate the implementation across tree sizes from 100 to 100{,}000 entries, measuring insert latency, proof generation time, proof verification time, and memory consumption. At 100{,}000 entries, the SMT achieves mean insert latency of 1.825\,ms (5.5$\times$ below the 10\,ms target), proof generation in 0.018\,ms (278$\times$ below target), and proof verification in 1.744\,ms (within the 2\,ms target). Insert latency is constant across tree sizes, confirming O($d$) complexity independent of entry count. A comparative benchmark demonstrates that Poseidon provides a 4.97$\times$ speedup over MiMC in native JavaScript execution, consistent with its lower R1CS constraint count (240 vs.\ 340 per hash). The implementation maintains full compatibility with the BN128 scalar field via circomlibjs, enabling direct integration with Circom/SnarkJS Groth16 proof circuits. These results validate the SMT as a viable foundation for enterprise ZK validium state management, with identified optimization paths for scaling beyond 100{,}000 entries.
\end{abstract}
