% discussion.tex -- Sparse Merkle Trees with Poseidon Hash

\section{Discussion}
\label{sec:discussion}

\subsection{Hypothesis Assessment}

The experimental hypothesis stated that a depth-32 SMT with Poseidon hash could support 100{,}000+ entries with insert latency $< 10$\,ms, proof generation $< 5$\,ms, and proof verification $< 2$\,ms in TypeScript, while maintaining BN128 field compatibility. All five criteria are met:

\begin{table}[htbp]
\centering
\caption{Hypothesis evaluation at $n = 100{,}000$ entries.}
\label{tab:hypothesis}
\begin{tabular}{llll}
\hline
\textbf{Target} & \textbf{Required} & \textbf{Measured (P95)} & \textbf{Margin} \\
\hline
Insert latency & $< 10$\,ms & 2.014\,ms & 5.0$\times$ \\
Proof generation & $< 5$\,ms & 0.021\,ms & 238$\times$ \\
Proof verification & $< 2$\,ms & 1.869\,ms & 1.07$\times$ \\
Memory & $< 2$\,GB & 233.9\,MB & 8.5$\times$ \\
BN128 compatibility & Full & Full & --- \\
\hline
\end{tabular}
\end{table}

The hypothesis is \textbf{confirmed}. However, the margins are not uniform, and the tightest constraint (proof verification at 1.07$\times$ P95 margin) warrants discussion.

\subsection{Trade-off Analysis}

\paragraph{Proof verification is the bottleneck.} Verification requires 32 sequential Poseidon hashes with no opportunity for parallelism or caching, making it the irreducible cost floor of the system. The P95 of 1.869\,ms leaves only 131\,$\mu$s of headroom against the 2\,ms target. On slower hardware or under concurrent load, this margin could vanish. Two mitigation paths exist: (1)~WebAssembly Poseidon implementations (e.g., circom-compat-wasm) offer approximately 10$\times$ speedup over JavaScript BigInt by using native 64-bit arithmetic, and (2)~a native Node.js addon in Rust or C++ could achieve 100$\times$ improvement.

\paragraph{Memory scales linearly.} At 100K entries, the implementation uses 233.9\,MB. Extrapolating the 17$\times$ node-to-entry ratio, 1M entries would require approximately 2.3\,GB, exceeding the 2\,GB threshold. For deployments beyond 100K entries, the in-memory \texttt{Map} should be replaced with a persistent key-value store (LevelDB or RocksDB) with an LRU cache for hot path nodes. This is an implementation concern, not an algorithmic one---the O($d$) complexity is unaffected.

\paragraph{Insert latency is dominated by hashing.} The 1.825\,ms mean insert latency comprises approximately 1.7\,ms of Poseidon computation (32 hashes) and approximately 0.1\,ms of hash map operations. This means insert performance cannot be significantly improved without changing the hash implementation. For the batch processing pipeline (where the SMT receives transactions in batches of 4--256), the batch update optimization of Chen et~al.~\cite{chen2023batch} can reduce amortized cost by sharing common path computation across multiple inserts.

\paragraph{Poseidon vs.\ MiMC: a settled question.} Our 4.97$\times$ native speedup, combined with the 29\% constraint reduction (240 vs.\ 340 R1CS), the production consensus (Polygon, Scroll, Iden3, Semaphore all use Poseidon), and identical BN128 field compatibility, leaves no technical argument for MiMC in this context. The only caveat is that MiMC may perform better under Plonkish arithmetization~\cite{guohanze2024benchmarking}, which is irrelevant to our Groth16/R1CS pipeline.

\subsection{Limitations}

\paragraph{Single-run, single-machine.} All benchmarks were conducted on a single Windows~11 workstation. Results may differ on Linux servers with different V8 JIT behavior, NUMA architectures, or containerized environments. The 50-sample methodology provides statistical confidence for this configuration but not generalizability.

\paragraph{Synthetic workload.} Keys are generated by a deterministic linear congruential generator, producing uniform distribution. Enterprise workloads may exhibit locality (e.g., sequential record IDs for a single enterprise), which could affect hash map performance. The tree structure itself is unaffected by key distribution due to the Poseidon key derivation, which distributes any input uniformly.

\paragraph{No concurrent access.} The implementation is single-threaded with no locking or transaction isolation. A production deployment processing concurrent transactions will require a concurrency control layer above the SMT.

\paragraph{JavaScript BigInt overhead.} Our measurements reflect JavaScript's arbitrary-precision BigInt, which is orders of magnitude slower than native field arithmetic in Rust or C++. The absolute numbers (1.8\,ms insert, 1.7\,ms verify) should be interpreted as upper bounds; a native implementation would likely achieve $10\times$--$100\times$ better performance.

\subsection{Reconciliation with Literature}

The 4.97$\times$ Poseidon-over-MiMC speedup in native execution is higher than the ${\sim}2\times$ proving time improvement reported by Semaphore~\cite{semaphore2024v4}. This is expected: Semaphore measures end-to-end circuit proving time (which includes witness generation, FFT, and multi-scalar multiplication in addition to hash computation), while we isolate native hash throughput. The ratio of constraint counts ($340/240 = 1.42\times$) represents a lower bound on the speedup; native execution amplifies this because MiMC's $x^7$ S-box requires more BigInt multiplications per round than Poseidon's $x^5$.

The 950$\times$ gap between our JavaScript insert latency and Rust (Monotree: 1.92\,$\mu$s) is consistent with known BigInt-to-native arithmetic overhead. The Reilabs LargeSMT~\cite{reilabs2024largesmt} multi-threaded numbers (63--193\,$\mu$s) narrow this to 9--29$\times$ when comparing single-threaded JavaScript against multi-threaded Rust, which remains consistent with the expected overhead range.
