% introduction.tex -- Sparse Merkle Trees with Poseidon Hash

\section{Introduction}
\label{sec:introduction}

Enterprise blockchain adoption remains constrained by a fundamental tension between transparency and privacy. Public blockchains provide verifiability and auditability but expose all transaction data, making them unsuitable for enterprises that must protect trade secrets, supplier relationships, and competitive pricing. Private blockchains preserve confidentiality but sacrifice the trust guarantees that make blockchain compelling in the first place. Zero-knowledge (ZK) proof systems offer a resolution to this tension: an enterprise can prove the correctness of its state transitions without revealing the underlying data~\cite{groth2016size}.

A critical architectural decision in any ZK-based enterprise system is the choice of state commitment structure. The state tree must satisfy three simultaneous requirements: (1)~it must support efficient membership and non-membership proofs for on-chain verification, (2)~it must operate natively over the finite field used by the ZK proof system to minimize circuit complexity, and (3)~it must scale to enterprise workloads---hundreds of thousands of state entries---with sub-second operation latency. These requirements are not independently satisfiable; the choice of hash function, tree structure, and field arithmetic are deeply coupled.

This paper presents the design, implementation, and empirical evaluation of a depth-32 Sparse Merkle Tree (SMT) using the Poseidon hash function~\cite{grassi2021poseidon} over the BN128 scalar field, purpose-built for the Basis Network enterprise ZK validium node. The Basis Network is an Avalanche Layer~1 blockchain serving Latin American enterprises, where each enterprise operates a validium node that maintains private state off-chain and submits ZK validity proofs to the L1. The SMT serves as the foundational state structure upon which all subsequent components---transaction queuing, batch aggregation, proof generation, and L1 submission---depend.

Our evaluation demonstrates that the proposed SMT implementation meets all target performance requirements with significant margin: insert latency of 1.825\,ms (5.5$\times$ below the 10\,ms target), proof generation in 0.018\,ms (278$\times$ below the 5\,ms target), and proof verification in 1.744\,ms (within the 2\,ms target), all measured at 100,000 entries on a standard development machine. We further establish that Poseidon provides a 4.97$\times$ speedup over MiMC in native JavaScript execution, consistent with the theoretical prediction based on relative constraint counts.
