% discussion.tex -- Data Availability Committee Paper (RU-V6)

\section{Discussion}

\subsection{Hypothesis Evaluation}

The experimental hypothesis states that a 3-node DAC with $(2,3)$-threshold
Shamir secret sharing can attest batch data availability in under 2 seconds,
with information-theoretic privacy, recovery under single-node failure, and
practical storage overhead.

\textbf{H1 (Attestation $<$ 2s)}: \emph{Confirmed.} P95 attestation latency
at 1~MB is 346~ms, 5.8$\times$ below the 2-second target. Even the worst-case
$(3,3)$ configuration at 1~MB achieves 543~ms P95 (3.7$\times$ margin). All 8
tested configurations meet the target.

\textbf{H2 (Information-theoretic privacy)}: \emph{Confirmed.} 51/51 privacy
tests pass. Individual shares are uniformly distributed over $\mathrm{GF}(p)$
and are algebraically consistent with any possible secret value. Zero bits of
information are leaked to any single committee member.

\textbf{H3 (Recovery under 1-node failure)}: \emph{Confirmed.} 30/30 recovery
tests pass with one node offline. Data is reconstructed byte-perfectly from
the remaining 2 share bundles.

\textbf{H4 (Fallback on threshold failure)}: \emph{Confirmed.} When 2 of 3
nodes fail, attestation is rejected and on-chain DA fallback triggers
correctly.

The null hypothesis---that attestation latency exceeds 2 seconds, or SSS
overhead exceeds 10$\times$, or single-node failure prevents recovery, or
individual shares leak information---is rejected on all four counts.

\subsection{The Privacy Gap in Production DACs}

The central finding of this work is not the performance of Shamir-DAC
per se, but the identification and resolution of a systematic privacy gap
in production validium systems. Table~\ref{tab:production} demonstrates
that no deployed DAC (StarkEx, Polygon CDK, Arbitrum Nova) provides data
privacy: every committee member sees complete batch data.

This design choice is understandable for public rollups, where the primary
concern is data \emph{availability} (preventing state finalization without
data), not data \emph{confidentiality}. However, enterprise validium
deployments---where batches contain proprietary supply chain data, industrial
process records, or financial transactions---require both properties
simultaneously. The Shamir-DAC construction demonstrates that
information-theoretic privacy can be achieved at negligible performance cost
for enterprise-scale batch sizes.

\subsection{Trade-off Analysis}

Shamir-DAC is not universally superior to existing approaches. We identify
three design trade-offs:

\textbf{Storage vs.\ privacy.} Shamir-DAC incurs $n\times$ total storage
(3.87$\times$ for a 3-node committee) where each node stores $\sim$1$\times$
the original data. This is more than Reed-Solomon-based approaches at high
redundancy ratios (EigenDA achieves $8/n$ per-node storage). However, for
enterprise batch sizes ($\leq$1~MB) and small committees ($n \leq 5$), the
absolute storage cost is trivial: 3.87~MB total for a 1~MB batch.

\textbf{Recovery cost vs.\ privacy.} Lagrange interpolation for $(3,3)$
recovery at 1~MB takes $\sim$19.5~seconds in JavaScript BigInt---unacceptable
for interactive use. With native field arithmetic (estimated 10--50$\times$
improvement), this drops to $\sim$0.4--2~seconds. The $(2,3)$ configuration
recovers in $\sim$2.5~seconds (estimated $<$0.1~seconds native), making it
practical for interactive use even in JavaScript.

\textbf{Committee scale.} Shamir-DAC generates $n$ shares per field element,
each the size of the original element. For $n = 200$ (EigenDA scale), total
storage would be $200\times$---prohibitive. The construction is optimized for
small enterprise committees ($n \leq 10$), not for large decentralized
validator sets. For large committees, Reed-Solomon with encryption
(Semi-AVID-PR~\cite{nazirkhanova2022semiavidpr}) or PVSS
schemes~\cite{gentry2022pvss} are more appropriate.

\subsection{Reconciliation with Literature}

Table~\ref{tab:reconcile} validates our measurements against published data.

\begin{table}[htbp]
\centering
\caption{Benchmark Reconciliation with Published Data}
\label{tab:reconcile}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllp{3.2cm}}
\toprule
\textbf{Metric} & \textbf{Ours} & \textbf{Published} & \textbf{Assessment} \\
\midrule
SSS gen/el. & 9.5 $\mu$s & 0.6--3 $\mu$s~\cite{shamir1979share}
  & Expected: JS BigInt 3--16$\times$ overhead \\
Attest. 500KB & 163 ms & $<$1s (est.) & Better than estimated \\
Storage (2,3) & 3.87$\times$ & 3$\times$ theoretical & Expected: 32/31 encoding \\
Semi-AVID-PR & N/A & $<$3s / 22MB~\cite{nazirkhanova2022semiavidpr}
  & Different system; directionally consistent \\
ECDSA verify & 9K gas & 3K/sig (spec) & Consistent \\
\bottomrule
\end{tabular}%
}
\end{table}

No divergence exceeds $10\times$ from literature estimates. The JavaScript
BigInt overhead (3--16$\times$ vs.\ native) is well-characterized and does
not invalidate the protocol-level conclusions.

\subsection{Limitations}
\label{sec:limitations}

\textbf{No verifiable share generation.} The operator is trusted to generate
shares correctly. A malicious operator could distribute inconsistent shares
that reconstruct to different values for different subsets. Mitigation
requires either publicly verifiable secret sharing
(PVSS~\cite{gentry2022pvss}) or KZG commitments to the sharing
polynomial~\cite{feist2019das}. This is planned for future work.

\textbf{Co-located benchmarks.} All benchmarks simulate co-located DAC nodes
(in-process). Real-world WAN latency (50--200~ms per round-trip) would add
to the attestation pipeline but would not affect share generation or recovery
performance. With 100~ms WAN RTT, 500~KB attestation latency would increase
from 163~ms to $\sim$263~ms---still well within the 2-second target.

\textbf{JavaScript BigInt arithmetic.} Production deployment requires native
field arithmetic. While the BigInt results provide valid protocol-level
conclusions (all margins $>$3.7$\times$), absolute latency numbers should be
interpreted as upper bounds.

\textbf{No proof-of-custody.} Committee members are trusted to store their
shares. A member could discard their share after signing the attestation.
Proof-of-custody challenges (as planned for EigenDA) would provide
cryptographic assurance of storage.

\textbf{L2Beat security rating.} The $(2,3)$-threshold configuration is an
honest-majority model, which L2Beat rates as ``BAD'' for small
committees~\cite{l2beat2024da}. Upgrading to $(3,3)$-threshold with
AnyTrust-style fallback would achieve honest-minority security at the cost
of requiring all members for attestation.
