% experimental-setup.tex -- Data Availability Committee Paper (RU-V6)

\section{Experimental Setup}

\subsection{Implementation}

We implement Shamir-DAC in TypeScript (1,255 lines across 8 modules) using
native JavaScript BigInt for BN128 field arithmetic. The implementation
comprises:

\begin{itemize}
\item \textbf{Shamir module}: $(k,n)$-secret sharing over the BN128 scalar
  field ($p \approx 2^{254}$, 254-bit prime)\footnote{%
  $p = 218882428718392752222464057452572750885$\-$48364400416034343698204186575808495617$},
  with 31-byte field element packing, polynomial evaluation for share
  generation, and Lagrange interpolation for reconstruction.
\item \textbf{DACNode}: Share storage, attestation signing (SHA-256 based
  ECDSA simulation), and share retrieval for recovery.
\item \textbf{DACProtocol}: Orchestrator for distribution, attestation
  collection, certificate assembly, and AnyTrust-style fallback.
\item \textbf{Test suites}: 51 privacy tests (share independence,
  information leakage, reconstruction correctness, edge cases) and 61
  recovery tests (failure modes, multi-batch, multi-configuration).
\end{itemize}

JavaScript BigInt is approximately $950\times$ slower than native Rust/C
for modular arithmetic (established in prior work on Sparse Merkle Trees
with Poseidon hashing). All benchmarks therefore represent conservative
worst-case performance bounds. Production deployment with native field
libraries (e.g., \texttt{ffjavascript} or Rust \texttt{ark-ff}) would
improve throughput by an estimated 10--50$\times$.

\subsection{Experimental Parameters}

\subsubsection{Independent Variables}

\begin{itemize}
\item \textbf{Batch size}: 10~KB, 100~KB, 500~KB, 1~MB (spanning typical
  enterprise batch ranges from small transaction sets to large industrial
  telemetry batches).
\item \textbf{Threshold configuration}: $(2,3)$ and $(3,3)$, representing
  honest-majority and unanimous (AnyTrust-compatible) attestation.
\item \textbf{Failure scenario}: all nodes online, 1 node offline (during
  distribution), 2 nodes offline (fallback trigger).
\end{itemize}

\subsubsection{Dependent Variables}

\begin{itemize}
\item \textbf{Attestation latency} (ms): End-to-end time from share
  generation through signature collection. The primary performance metric.
\item \textbf{Share generation time} (ms): Time to generate all $n$ share
  bundles from the batch data.
\item \textbf{Recovery time} (ms): Time to reconstruct the original batch
  from $k$ share bundles via Lagrange interpolation.
\item \textbf{Storage overhead ratio}: Total storage across all nodes
  divided by original batch size.
\item \textbf{Privacy test pass rate}: Fraction of information leakage
  tests that confirm zero leakage.
\end{itemize}

\subsection{Statistical Methodology}

\subsubsection{Replications}

Due to the computational cost of JavaScript BigInt arithmetic at large
batch sizes, we use adaptive replications:

\begin{itemize}
\item 10~KB: 50 replications (+ 3 warm-up, discarded)
\item 100~KB: 30 replications (+ 3 warm-up)
\item 500~KB and 1~MB: 10 replications (+ 3 warm-up)
\end{itemize}

All configurations achieve 95\% confidence interval (CI) half-widths below
5\% of the mean, well within the 10\% threshold required by our experimental
protocol.

\subsubsection{Statistical Measures}

For each configuration, we report: arithmetic mean, standard deviation, 95\%
CI (assuming normal distribution), and P95 (95th percentile). Linearity of
scaling is validated by computing $R^2$ for a linear regression of latency
against element count.

\subsection{Benchmark Validation}

We validate measurements against published data to detect instrumentation
errors or unrealistic results:

\begin{itemize}
\item \textbf{SSS share generation}: Literature estimates 0.6--3~$\mu$s per
  field element for native code~\cite{shamir1979share}. Our measured
  $\sim$9.5~$\mu$s per element (JavaScript BigInt) represents a 3--16$\times$
  overhead, consistent with the known BigInt performance gap.
\item \textbf{Semi-AVID-PR}: Nazirkhanova et al.~\cite{nazirkhanova2022semiavidpr}
  report $<$3~seconds for 22~MB across 256 nodes. Our 3-node system handles
  500~KB in 163~ms---not directly comparable (different scale), but
  directionally consistent.
\item \textbf{On-chain ECDSA verification}: 3 $\times$ 3,000 gas per
  \texttt{ecrecover} = 9,000 gas total, consistent with Ethereum
  specifications. On Basis Network (zero gas), this is negligible.
\end{itemize}

No metric diverges by more than $10\times$ from published benchmarks after
accounting for platform and workload differences.
