% results.tex -- Data Availability Committee Paper (RU-V6)

\section{Results}

All experiments were conducted on Node.js v22.13.1, Windows 11, commodity
desktop hardware. Results represent worst-case JavaScript BigInt performance.

\subsection{Attestation Pipeline Latency}

Table~\ref{tab:attestation} presents the attestation pipeline latency (share
generation + distribution + signature collection) across batch sizes and
threshold configurations. This is the latency-critical metric: the time
between the operator producing a batch and receiving a valid DAC certificate.

\begin{table}[h]
\centering
\caption{Attestation Pipeline Latency by Configuration}
\label{tab:attestation}
\begin{tabular}{llrrrr}
\toprule
\textbf{Config} & \textbf{Batch} & \textbf{Mean} & \textbf{P95} &
\textbf{Target} & \textbf{Margin} \\
 & \textbf{Size} & \textbf{(ms)} & \textbf{(ms)} & \textbf{(ms)} & \\
\midrule
2-of-3 & 10 KB  & 3.2   & 3.9   & 2,000 & 512$\times$ \\
2-of-3 & 100 KB & 32.1  & 34.4  & 2,000 & 58$\times$ \\
2-of-3 & 500 KB & 163.5 & 175.3 & 2,000 & 11$\times$ \\
2-of-3 & 1 MB   & 320.2 & 346.2 & 2,000 & 5.8$\times$ \\
\midrule
3-of-3 & 10 KB  & 5.3   & 7.1   & 2,000 & 282$\times$ \\
3-of-3 & 100 KB & 52.6  & 54.8  & 2,000 & 37$\times$ \\
3-of-3 & 500 KB & 269.4 & 284.6 & 2,000 & 7.0$\times$ \\
3-of-3 & 1 MB   & 526.2 & 542.5 & 2,000 & 3.7$\times$ \\
\midrule
2-of-3$^*$ & 100 KB & 30.9 & 36.0 & 2,000 & 56$\times$ \\
\bottomrule
\end{tabular}
\begin{flushleft}
\small $^*$One node offline during distribution (degraded mode).
\end{flushleft}
\end{table}

All configurations meet the 2-second target with at least 3.7$\times$ margin.
Attestation latency scales linearly with data size at approximately
0.32~ms/KB for $(2,3)$ and 0.52~ms/KB for $(3,3)$. The $(3,3)$ configuration
is $\sim$1.6$\times$ slower than $(2,3)$ due to additional polynomial
evaluations per share (degree 2 vs.\ degree 1). Degraded-mode operation (1
node offline) shows no statistically significant latency increase over
normal operation.

\subsection{Share Generation Breakdown}

Table~\ref{tab:sharegen} isolates the share generation cost, which dominates
the attestation pipeline.

\begin{table}[h]
\centering
\caption{Share Generation Latency (2-of-3 Configuration)}
\label{tab:sharegen}
\begin{tabular}{rrrr}
\toprule
\textbf{Batch Size} & \textbf{Elements} & \textbf{Mean (ms)} &
\textbf{Per-Element ($\mu$s)} \\
\midrule
10 KB  & 323    & 3.1   & 9.6  \\
100 KB & 3,226  & 31.1  & 9.6  \\
500 KB & 16,130 & 163.2 & 10.1 \\
1 MB   & 32,259 & 318.1 & 9.9  \\
\bottomrule
\end{tabular}
\end{table}

Per-element cost is constant at $\sim$9.5--10.0~$\mu$s (JavaScript BigInt),
confirming $O(kn \cdot m)$ linear scaling. Linear regression yields
$R^2 > 0.999$. Literature estimates for native implementations are
0.6--3~$\mu$s per element; the 3--16$\times$ overhead is consistent with
JavaScript BigInt's known performance characteristics. Estimated production
per-element cost with native field arithmetic: $<$1~$\mu$s.

\subsection{Data Recovery Latency}

Table~\ref{tab:recovery} presents the recovery latency for reconstructing
the original batch from $k$ share bundles.

\begin{table}[h]
\centering
\caption{Data Recovery Latency by Configuration}
\label{tab:recovery}
\begin{tabular}{llrrr}
\toprule
\textbf{Config} & \textbf{Batch} & \textbf{Elements} & \textbf{Mean (ms)} &
\textbf{Per-El. ($\mu$s)} \\
\midrule
2-of-3 & 10 KB  & 323    & 25.0     & 77.4  \\
2-of-3 & 100 KB & 3,226  & 250.6    & 77.7  \\
2-of-3 & 500 KB & 16,130 & 1,260.3  & 78.1  \\
2-of-3 & 1 MB   & 32,259 & 2,482.1  & 76.9  \\
\midrule
3-of-3 & 10 KB  & 323    & 195.7    & 606.0 \\
3-of-3 & 100 KB & 3,226  & 1,946.3  & 603.3 \\
3-of-3 & 500 KB & 16,130 & 9,745.8  & 604.2 \\
3-of-3 & 1 MB   & 32,259 & 19,478.8 & 603.7 \\
\bottomrule
\end{tabular}
\end{table}

Recovery is $\sim$8$\times$ slower for $(3,3)$ than $(2,3)$, reflecting the
$O(k^2)$ complexity of Lagrange interpolation: $(3,3)$ requires 9 field
multiplications per element vs.\ $(2,3)$'s 4. At 1~MB, $(2,3)$ recovery
takes $\sim$2.5~seconds; $(3,3)$ takes $\sim$19.5~seconds.

Critically, \emph{recovery is not on the attestation critical path}. Recovery
is invoked only when data reconstruction is explicitly requested (e.g.,
enterprise withdrawal proof generation, auditor data verification). The
attestation latency in Table~\ref{tab:attestation} determines system
throughput; recovery latency affects only interactive data retrieval.

\subsection{Storage Overhead}

\begin{table}[h]
\centering
\caption{Storage Overhead (3-Node Committee)}
\label{tab:storage}
\begin{tabular}{rrrr}
\toprule
\textbf{Batch Size} & \textbf{Per-Node (bytes)} & \textbf{Total (bytes)} &
\textbf{Overhead} \\
\midrule
10 KB  & 12,984    & 38,952    & 3.91$\times$ \\
100 KB & 129,104   & 387,312   & 3.88$\times$ \\
500 KB & 645,264   & 1,935,792 & 3.87$\times$ \\
1 MB   & 1,290,416 & 3,871,248 & 3.87$\times$ \\
\bottomrule
\end{tabular}
\end{table}

Overhead converges to 3.87$\times$ at scale, explained by $n \times 32/31$:
each field element encodes 31 bytes but is stored as 32 bytes. Each node
stores approximately $1\times$ the original data size. Table~\ref{tab:overhead_compare}
compares with production systems.

\begin{table}[h]
\centering
\caption{Storage Overhead Comparison with Production Systems}
\label{tab:overhead_compare}
\begin{tabular}{lrrl}
\toprule
\textbf{System} & \textbf{Total} & \textbf{Per-Node} &
\textbf{Recovery From} \\
\midrule
Full replication (StarkEx) & $n \times$ & $1\times$ & Any 1 node \\
Semi-AVID-PR               & 3.2$\times$ & $3.2/n$ & Any $n/3$ nodes \\
\textbf{Shamir-DAC (ours)} & \textbf{3.87$\times$} & $\sim$\textbf{1$\times$} & \textbf{Any $k$ nodes} \\
Celestia (2D RS)           & 4$\times$   & $4/n$   & Any $n/4$ nodes \\
EigenDA V2 (RS 1/8)        & 8$\times$   & $8/n$   & Any $n/8$ nodes \\
\bottomrule
\end{tabular}
\end{table}

Shamir-DAC's 3.87$\times$ overhead is less than EigenDA (8$\times$),
comparable to Celestia (4$\times$) and Semi-AVID-PR (3.2$\times$), and
marginally above full replication (3$\times$ for 3 nodes). The additional
0.87$\times$ over full replication is the 32/31 encoding overhead---the
cost of information-theoretic privacy.

\subsection{Privacy Verification}

51 of 51 privacy tests pass. The test suite validates:

\begin{itemize}
\item \textbf{Reconstruction correctness}: All $\binom{n}{k}$ subsets of $k$
  shares reconstruct the original data byte-perfectly.
\item \textbf{Share independence}: Share distributions for different secrets
  are statistically indistinguishable ($|\Delta| = 3$--5\% with $n=1{,}000$
  samples, consistent with sampling variance).
\item \textbf{$k-1$ information leakage}: A single share is algebraically
  consistent with any candidate secret value (verified for 100 candidate
  secrets per test).
\item \textbf{Edge cases}: Zero-length data, all-0xFF data, secret $= 0$,
  secret $= p-1$, single-byte data.
\end{itemize}

\subsection{Failure Mode Verification}

61 of 61 recovery and failure mode tests pass:

\begin{table}[h]
\centering
\caption{Failure Mode Test Results}
\label{tab:failure}
\begin{tabular}{lrrl}
\toprule
\textbf{Scenario} & \textbf{Tests} & \textbf{Passed} & \textbf{Outcome} \\
\midrule
All 3 online            & 10 & 10 & 3/3 signatures \\
1 node offline          & 30 & 30 & 2/3 valid, recovery OK \\
2 nodes offline         & 5  & 5  & Fallback triggered \\
Node offline mid-distrib. & 5 & 5 & 2/3 receive, valid \\
Node rejoin             & 5  & 5  & Full attestation resumes \\
Certificate tampered    & 3  & 3  & Detected, rejected \\
20 sequential batches   & 1  & 1  & All recoverable \\
3-of-3 configuration   & 1  & 1  & Correct behavior \\
3-of-5 configuration   & 1  & 1  & Correct behavior \\
\bottomrule
\end{tabular}
\end{table}

The $(2,3)$-threshold configuration tolerates exactly 1 node failure for both
attestation and recovery, as designed. When 2 of 3 nodes fail, the attestation
threshold cannot be met and fallback to on-chain DA triggers correctly.
