% system-model.tex -- Data Availability Committee Paper (RU-V6)

\section{System Model}

\subsection{Network Model}

The system comprises an \emph{enterprise validium node} (the operator) and a
\emph{data availability committee} of $n$ nodes. The operator receives
transactions from enterprise systems (e.g., industrial traceability, ERP),
maintains an off-chain Sparse Merkle Tree of enterprise state, aggregates
transactions into batches, generates Groth16 ZK proofs~\cite{groth2016size},
and submits proofs with state roots to the Basis Network L1 (an Avalanche
Subnet-EVM chain with zero-fee gas model).

Communication between the operator and DAC nodes is over authenticated
channels. We assume eventual delivery: messages may be delayed but are not
permanently lost between non-faulty nodes. DAC nodes maintain persistent
storage for their assigned shares.

\subsection{Trust Assumptions}

We adopt a $(k,n)$-threshold trust model where:

\begin{itemize}
\item At least $k$ of $n$ committee members are honest (store their shares
  faithfully and respond to attestation and recovery requests).
\item Honest members do not collude to reconstruct data outside the
  protocol-specified recovery procedure.
\item The operator is trusted to generate shares correctly (verifiable share
  generation is a non-goal for the MVP; see Section~\ref{sec:limitations}).
\end{itemize}

For the enterprise deployment target, $n = 3$ and $k = 2$: any 2 of 3 members
can attest data availability and reconstruct the original batch. This is an
honest-majority model. Following the AnyTrust design
pattern~\cite{arbitrum2024anytrust}, the system includes a liveness fallback:
if fewer than $k$ members are available for attestation, the operator posts
batch data directly on-chain, temporarily converting the system from validium
to rollup mode.

\subsection{Enterprise Threat Model}

We consider four classes of adversary:

\textbf{Curious Committee Member.} A single DAC member attempts to learn
the contents of a batch from their stored share alone. Under $(k,n)$-Shamir
secret sharing with $k \geq 2$, a single share is uniformly distributed over
$\mathrm{GF}(p)$ and is consistent with \emph{every possible secret}
value~\cite{shamir1979share}. The information leakage is exactly 0 bits---this
holds against computationally unbounded adversaries.

\textbf{Colluding Minority.} Up to $k-1$ committee members pool their shares
to learn batch contents. Under Shamir's scheme, $k-1$ shares are jointly
uniformly distributed and reveal zero information about the secret. For
$(2,3)$: a single compromised member learns nothing. For $(3,5)$: two
colluding members learn nothing.

\textbf{Data Withholding.} Committee members sign the attestation certificate
without actually storing their shares. This is the classical DAC trust
assumption~\cite{ethereum2024validium}. Mitigation requires either
proof-of-custody challenges (future work) or reputational/contractual
enforcement.

\textbf{Operator Collusion.} The operator colludes with $k$ or more committee
members to finalize state with unavailable data. The ZK proof guarantees state
transition \emph{validity} (the new root is correctly derived from the old
root and the batch), but data \emph{availability} is compromised: users cannot
reconstruct state to generate withdrawal proofs. This is the fundamental
validium trust assumption, shared by all systems in Table~\ref{tab:production}.

\subsection{Security Properties}

We formalize five invariants that Shamir-DAC must satisfy:

\begin{enumerate}
\item[\textbf{INV-DA1}] \textbf{Privacy.} No individual DAC member can
  reconstruct the complete batch data from their share alone.
\item[\textbf{INV-DA2}] \textbf{Availability.} If at least $k$ of $n$ members
  respond, the complete batch data can be reconstructed byte-perfectly.
\item[\textbf{INV-DA3}] \textbf{Attestation Soundness.} A batch is attested as
  available on-chain only if at least $m$ of $n$ members have signed the data
  commitment hash ($m$ = attestation threshold, aligned with $k$).
\item[\textbf{INV-DA4}] \textbf{Liveness Fallback.} If fewer than $m$ members
  are available, the system falls back to on-chain DA within a bounded timeout.
\item[\textbf{INV-DA5}] \textbf{Commitment Binding.} The data commitment hash
  committed on-chain binds to the exact batch data; no two distinct batches
  produce the same commitment (collision resistance of SHA-256).
\end{enumerate}
