% system-model.tex -- Cross-Enterprise Verification Paper (RU-V7)

\section{System Model}

\subsection{Network Architecture}

We consider a hub-and-spoke validium architecture deployed on an Avalanche L1
(Subnet-EVM) with zero gas fees and permissioned access. The system comprises:

\begin{itemize}
\item \textbf{Enterprise nodes} ($E_1, \ldots, E_N$): Each enterprise
  operates an independent validium node maintaining a depth-32 sparse Merkle
  tree (SMT) over the BN128 scalar field. Enterprises process private
  transactions, aggregate them into batches, generate Groth16 proofs of
  correct state transitions, and submit proofs to the L1.

\item \textbf{Settlement layer} (Basis Network L1): An Avalanche Subnet-EVM
  chain (Chain ID 43199) that verifies ZK proofs and stores state roots for
  each enterprise. The L1 is the single source of truth for state root
  progression.

\item \textbf{Hub coordinator}: A logical entity (implemented as an L1
  contract or off-chain service) that collects cross-reference proofs from
  enterprise pairs and, optionally, batches them for aggregated verification.
\end{itemize}

\subsection{Trust Assumptions}

\begin{enumerate}
\item \textbf{Proof system soundness}: Groth16 proofs on BN128 are
  computationally sound under the $q$-type assumption in the generic bilinear
  group model~\cite{groth2016size}. An adversary cannot produce a valid proof
  for a false statement except with negligible probability.

\item \textbf{Hash function security}: Poseidon over BN128~\cite{grassi2021poseidon}
  provides 128-bit collision resistance and preimage resistance. The
  interaction commitment $c = \text{Poseidon}(k_A, v_A, k_B, v_B)$ is
  computationally binding and hiding.

\item \textbf{Enterprise isolation}: Each enterprise's SMT is maintained
  independently. Enterprise $A$ does not have access to Enterprise $B$'s tree,
  keys, values, or Merkle proofs.

\item \textbf{State root integrity}: State roots submitted to the L1 are
  verified by ZK proofs. A malicious enterprise cannot submit a false state
  root without breaking Groth16 soundness.

\item \textbf{Hub coordinator honesty}: The hub coordinator is \emph{not}
  trusted with data. In the batched pairing approach, the coordinator receives
  only Groth16 proofs and public inputs (state roots and interaction
  commitments)---it never sees private keys, values, or Merkle paths.
\end{enumerate}

\subsection{Threat Model}

We consider three adversary classes:

\textbf{Passive observer}: Has access to all L1 transactions and public
signals. Can see state roots, interaction commitments, and proof submission
timestamps. Cannot break Poseidon preimage resistance (128-bit). Learns 1 bit
per cross-reference submission: the existence of an interaction between two
identified enterprises.

\textbf{Malicious enterprise}: Enterprise $A$ attempts to learn Enterprise
$B$'s private data by manipulating its own state or crafting adversarial
cross-reference proofs. Cannot succeed because: (a) the ZK proof system
ensures $A$ cannot produce a valid cross-reference proof without a valid
Merkle path in $B$'s tree, and (b) $A$ does not have access to $B$'s SMT.

\textbf{Compromised hub coordinator}: The coordinator receives proofs and
public inputs but not private witness data. Even if fully compromised, the
coordinator learns only what a passive observer learns: state roots (already
public) and interaction commitments (Poseidon hashes revealing no content).

\subsection{Privacy Definition}

We define cross-enterprise privacy as follows:

\begin{definition}[Cross-Enterprise Privacy]
A cross-enterprise verification protocol satisfies \emph{interaction privacy}
if, for any two valid cross-reference proofs $\pi_1$ and $\pi_2$ with the
same public inputs $(r_A, r_B, c_1)$ and $(r_A, r_B, c_2)$ where $c_1 \neq
c_2$, no probabilistic polynomial-time adversary can determine the private
witness $(k_A, v_A, k_B, v_B)$ from the public signals with probability
non-negligibly greater than random guessing.
\end{definition}

This follows directly from the zero-knowledge property of Groth16 and the
preimage resistance of Poseidon. The \emph{unavoidable} leakage is 1 bit per
interaction: the fact that a cross-reference proof was submitted, which
inherently reveals that an interaction exists between the two enterprises
identified by their state roots.
