% reproducibility.tex -- Batch Aggregation Paper (BNR-2026-004)

\section{Reproducibility}
\label{sec:reproducibility}

This section provides instructions for replicating the TLA+ model checking
results, including both the NoLoss counterexample discovery (v0) and the
corrected protocol verification (v1-fix).

\subsection{Prerequisites}

\begin{itemize}
\item \textbf{TLA+ Toolbox} or standalone TLC: Version 2.16 or later.
  Available from \url{https://github.com/tlaplus/tlaplus/releases}.
\item \textbf{Java}: JDK 8 or later (required by TLC).
\item Specifications are located in the repository under:\\
  \texttt{validium/specs/units/2026-03-batch-aggregation/}
\end{itemize}

\subsection{Replicating the NoLoss Counterexample (v0)}

The original specification with the premature checkpoint bug is in
\texttt{1-formalization/v0-analysis/}. To reproduce the counterexample:

\begin{enumerate}
\item Navigate to the experiment directory:\\
  \texttt{v0-analysis/experiments/BatchAggregation/}
\item The model configuration (\texttt{MC\_BatchAggregation.cfg}) uses
  10 transaction identifiers (\texttt{tx1}--\texttt{tx10}),
  \texttt{BatchSizeThreshold~=~4}, and breadth-first search.
\item Run TLC:
\end{enumerate}

\begin{quote}
\small
\texttt{java -jar tla2tools.jar -config MC\_BatchAggregation.cfg}\\
\texttt{~~~~MC\_BatchAggregation.tla}
\end{quote}

\textbf{Expected output}: TLC reports \texttt{Invariant NoLoss is violated}
with a 6-state counterexample trace: Init $\rightarrow$ Enqueue(tx1)
$\rightarrow$ Enqueue(tx2) $\rightarrow$ TimerTick $\rightarrow$ FormBatch
$\rightarrow$ Crash. The model checker generates 3,262 states (2,899
distinct) and terminates in under one second. The counterexample
demonstrates that \texttt{checkpointSeq} advances to 2 at batch formation,
so after crash recovery, tx1 and tx2 are irrecoverably lost.

\subsection{Verifying the Corrected Protocol (v1-fix)}

The corrected specification with deferred checkpoint is in
\texttt{1-formalization/v1-fix/}. To verify all properties:

\begin{enumerate}
\item Navigate to the experiment directory:\\
  \texttt{v1-fix/experiments/BatchAggregation/}
\item The model configuration uses 4 transaction identifiers
  (\texttt{tx1}--\texttt{tx4}), \texttt{BatchSizeThreshold~=~2},
  and breadth-first search with temporal property checking.
\item Run TLC:
\end{enumerate}

\begin{quote}
\small
\texttt{java -jar tla2tools.jar -config MC\_BatchAggregation.cfg}\\
\texttt{~~~~MC\_BatchAggregation.tla}
\end{quote}

\textbf{Expected output}: \texttt{Model checking completed. No error has
been found.} TLC generates 6,763 states (2,630 distinct), explores the
complete state graph to depth 18, and verifies all 6 safety invariants
(\texttt{TypeOK}, \texttt{NoLoss}, \texttt{NoDuplication},
\texttt{QueueWalConsistency}, \texttt{FIFOOrdering},
\texttt{BatchSizeBound}) and 1 liveness property
(\texttt{EventualProcessing}) with zero errors. Temporal property checking
completes in under one second.

\subsection{Replicating Empirical Benchmarks}

The benchmark code is located in
\texttt{validium/research/experiments/2026-03-18\_batch-aggregation/code/}.
Requirements: Node.js v22+ and TypeScript. Run:

\begin{quote}
\small
\texttt{npx ts-node src/benchmark.ts}\\
\texttt{npx ts-node src/test-determinism.ts}\\
\texttt{npx ts-node src/test-crash-recovery.ts}
\end{quote}

Results are written to the \texttt{results/} directory as JSON files.
Benchmark conditions: 30 replications per configuration, 2 warmup batches
discarded, Poisson-like arrival jitter. Platform-dependent metrics (WAL
write latency, fsync latency) may vary across operating systems and
storage hardware.
