% experimental-setup.tex -- Cross-Enterprise Verification Paper (RU-V7)

\section{Experimental Setup}

\subsection{Implementation}

The benchmark implementation is written in TypeScript using
circomlibjs~\cite{circomlibjs2024} for Poseidon hashing over the BN128 scalar
field ($p = 2^{254} + \epsilon$). The sparse Merkle tree implementation follows
the design validated in RU-V1~\cite{ruv1smt}, with depth 32 and Poseidon-2
internal nodes.

Gas costs are computed analytically using the EVM gas schedule (Cancun fork),
calibrated against on-chain measurements from RU-V3~\cite{ruv3gas}. The
analytical model accounts for:
\begin{itemize}
\item \texttt{ecPairing} precompile: 45,000 base + 34,000 per pair
\item MSM for public inputs: 6,150 per input + ecMul/ecAdd operations
\item Storage operations: SSTORE (22,100 cold / 5,000 warm), SLOAD (2,100
  cold / 100 warm)
\item Event emission: 375 base + 8 per byte
\item Calldata: 16 per nonzero byte, 4 per zero byte
\end{itemize}

The baseline single-enterprise gas cost ($285{,}756$ gas) was measured on
Subnet-EVM in RU-V3 and used as the reference for overhead ratio computation.

\subsection{Enterprise Configuration}

Five enterprises are configured with populated SMTs:
\begin{itemize}
\item Enterprises A, B: 100 entries each (simulating active enterprises)
\item Enterprises C, D: 50 entries each
\item Enterprise E: 25 entries
\end{itemize}

Cross-enterprise interactions are simulated by inserting matching records in
two enterprises' trees: Enterprise $A$ records an outgoing transaction and
Enterprise $B$ records the corresponding incoming transaction, both with the
same value.

\subsection{Parameters Varied}

\begin{itemize}
\item \textbf{Enterprise count} ($N$): 2, 3, 5, 10, 20, 50
\item \textbf{Interaction count} ($M$): Linear ($M = N-1$) and dense
  ($M > N$) configurations
\item \textbf{Verification approach}: Sequential, Batched Pairing, Hub
  Aggregation
\end{itemize}

\subsection{Metrics}

\begin{itemize}
\item \textbf{Total gas}: L1 gas cost for verifying all enterprise proofs
  plus cross-reference proofs, including storage and event emissions
\item \textbf{Overhead ratio}: Total gas divided by baseline gas (individual
  enterprise verification only, $N \times 285{,}756$)
\item \textbf{Cross-reference constraint count}: Number of R1CS constraints
  in the cross-reference circuit
\item \textbf{Proving time}: Estimated from the constraint-to-time ratio
  established in RU-V2~\cite{ruv2circuit} (65~$\mu$s/constraint for snarkjs,
  6.5~$\mu$s/constraint for rapidsnark)
\item \textbf{Privacy leakage}: Bits of information revealed beyond what is
  already public
\item \textbf{Merkle proof generation time}: Wall-clock time for SMT proof
  extraction
\item \textbf{Witness generation time}: Wall-clock time for cross-reference
  witness computation
\end{itemize}

\subsection{Statistical Methodology}

Timing benchmarks (Merkle proof generation, witness generation, verification
simulation) use 50 repetitions with independent invocations. We report mean
and standard deviation. Gas costs are deterministic (analytical model) and
require no statistical treatment.

\subsection{Benchmark Validation}

To ensure our analytical gas model is accurate, we validate against published
benchmarks and our own on-chain measurements:

\begin{itemize}
\item Groth16 verification gas (4 inputs): Our model yields 205,600 gas;
  Nebra~\cite{nebra2024upa} reports $\sim$205K gas for the same configuration.
  Ratio: 1.00$\times$.
\item SMT proof generation (depth 32): 0.018~ms mean; RU-V1 reported 0.02~ms.
  Ratio: 0.9$\times$.
\item Verification simulation (2 paths): 3.46~ms; RU-V1 reported 1.7~ms per
  path ($2 \times 1.7 = 3.4$~ms expected). Ratio: 1.02$\times$.
\item Constraint count formula: $1{,}038 \times (d+1)$ per path, validated
  in RU-V2. Our depth-32 path yields 34,254, matching exactly.
\end{itemize}

All metrics are directionally consistent with published benchmarks with no
divergence exceeding 1.1$\times$.
