% abstract.tex -- BNR-2026-013

\begin{abstract}
Enterprise zero-knowledge (ZK) rollups on Avalanche require proof
systems that balance verification cost, proof size, proving time, and
operational flexibility. The Basis Network Validium MVP currently
employs Groth16 via Circom/SnarkJS, achieving 306K gas verification
at 128-byte proof size but incurring per-circuit trusted setup
ceremonies that impede iterative circuit development. This paper
evaluates the migration from Groth16 to halo2-KZG (PLONKish
arithmetization with KZG polynomial commitments on BN254) for the
Basis Network zkEVM L2 prover. We conduct a systematic comparison
across five candidate proof systems---Groth16, PLONK-KZG, halo2-KZG,
halo2-IPA, and plonky2---eliminating candidates against enterprise
constraints (verification gas $< 500$K, proof size $< 1$~KB, BN254
EVM compatibility). We implement Rust benchmarks comparing Groth16
(arkworks) and halo2-KZG (PSE fork) across arithmetic chain and hash
chain circuits at varying scales (10--500 steps, 30 iterations each).
Results show that halo2-KZG proving time converges to within 1.2$\times$
of Groth16 at 500-step circuits, with custom degree-5 gates reducing
PLONKish rows by 2.4$\times$ for hash operations and a projected
17$\times$ for full Poseidon hashing. Proof sizes (672--800~bytes) and
published verification gas (290--420K) satisfy all targets.
We recommend halo2-KZG (Axiom fork) with universal Structured
Reference String as the production proof system for Basis Network
zkEVM L2, with a dual-verification migration strategy to preserve
backward compatibility during the transition period.
\end{abstract}
