% conclusion.tex -- BNR-2026-013

\section{Conclusion}
\label{sec:conclusion}

This paper evaluated the migration from Groth16 to PLONKish proof
systems for the Basis Network enterprise zkEVM L2. Through
systematic elimination of five candidates against enterprise
constraints and Rust benchmarks comparing Groth16 (arkworks)
and halo2-KZG (PSE fork) across arithmetic and hash chain
circuits, we established the following:

\begin{enumerate}
  \item \textbf{halo2-KZG satisfies all enterprise constraints}:
    verification gas 290--420K ($< 500$K), proof size 672--800~bytes
    ($< 1$~KB), proving time converging to 1.2$\times$ of Groth16
    at scale, universal SRS, custom gate support, native recursion,
    and Rust-native implementation.

  \item \textbf{Custom gates provide significant constraint
    reduction}: 2.4$\times$ measured for degree-5 S-box gates,
    17$\times$ projected for full Poseidon hashing, and
    100--900$\times$ for individual EVM operations. For
    Poseidon-heavy enterprise circuits, this translates to a
    projected 5--9$\times$ reduction in circuit row count.

  \item \textbf{Proving time converges at production scale}: The
    halo2-KZG overhead of 3--5$\times$ at small circuit sizes
    reduces to 1.2$\times$ at 500~steps, with asymptotic parity
    projected at production scale (30K--50K rows). Combined with
    custom gate row reduction, proving time is projected to
    improve over the current Groth16 baseline.

  \item \textbf{Universal SRS eliminates operational friction}: A
    single Powers-of-Tau ceremony replaces per-circuit Phase~2
    ceremonies, enabling rapid circuit iteration during zkEVM
    development.
\end{enumerate}

We recommend \textbf{halo2-KZG via the Axiom fork} as the
production proof system for Basis Network zkEVM L2, with:

\begin{itemize}
  \item Library: \texttt{axiom-crypto/halo2} (active development,
    KZG on BN254, production-proven on Ethereum mainnet)
  \item SRS: PSE perpetual-powers-of-tau, degree $k = 20$
    ($\sim$1M rows)
  \item Custom gates: ArithmeticGate, PoseidonGate, BitwiseGate,
    StorageGate
  \item Migration: Dual-verification router during transition
    period
\end{itemize}

This recommendation feeds downstream to the Logicist (TLA+
formalization of soundness preservation and migration safety),
the Architect (Rust implementation of the halo2-KZG prover and
PLONKVerifier contract), and the Prover (Coq verification of
custom gate correctness and proof system isomorphism).

\subsection{Future Work}

Five directions remain for subsequent research units:

\begin{enumerate}
  \item Optimal SRS degree ($k$) selection for enterprise circuits
    with varying batch sizes.
  \item GPU acceleration via ICICLE backend for halo2-KZG proving.
  \item FFLONK wrapper for sub-200K gas verification (optimization,
    not required for initial deployment).
  \item Proof aggregation via snark-verifier for multi-enterprise
    batching (RU-L10).
  \item Poseidon2 custom gate evaluation vs original Poseidon for
    further round count reduction.
\end{enumerate}
