% system-model.tex -- BNR-2026-013

\section{System Model}
\label{sec:system}

\subsection{Basis Network Architecture}

Basis Network operates as an Avalanche L1 (Subnet-EVM) with
Chain~ID~43199, providing zero-fee permissioned infrastructure
for enterprise traceability. The current Validium MVP processes
enterprise transactions through an off-chain node that:

\begin{enumerate}
  \item Maintains a Sparse Merkle Tree (depth~32) of enterprise
    state using Poseidon hashing over BN254.
  \item Aggregates transactions into batches (currently batch
    size~8).
  \item Generates Groth16 proofs via Circom/SnarkJS (274,291
    R1CS constraints, 12.9~s proving time).
  \item Submits proofs and state roots to the L1 for on-chain
    verification (306K gas via delegated Groth16Verifier).
\end{enumerate}

The evolution to a zkEVM L2 requires each enterprise to deploy
its own chain with full EVM compatibility, necessitating a more
flexible proof system.

\subsection{Enterprise Constraints}

The proof system must satisfy the invariants in
Table~\ref{tab:constraints}, derived from the Basis Network
objectives and invariants specification (ZK-01).

\begin{table}[htbp]
\caption{Enterprise Proof System Constraints}
\label{tab:constraints}
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{ID} & \textbf{Constraint} & \textbf{Target} \\
\midrule
C1 & Verification gas & $< 500$K \\
C2 & Proof size & $< 1$~KB \\
C3 & Proving time ratio vs Groth16 & $< 2\times$ at scale \\
C4 & Setup type & Universal or transparent \\
C5 & EVM curve compatibility & BN254 (EIP-196/197) \\
C6 & Custom gate support & Required \\
C7 & Recursive proof support & Required \\
C8 & Implementation language & Rust-native \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Candidate Proof Systems}

We evaluate five proof systems spanning three commitment
schemes:

\begin{itemize}
  \item \textbf{Groth16} (BN254): R1CS arithmetization, per-circuit
    trusted setup, pairing-based verification. Current production
    system.
  \item \textbf{PLONK-KZG} (BN254): PLONKish arithmetization,
    universal KZG-based SRS, pairing-based verification.
  \item \textbf{halo2-KZG} (BN254): PLONKish arithmetization via
    halo2 framework, KZG commitments on BN254, custom gates and
    lookup tables.
  \item \textbf{halo2-IPA} (Pasta): PLONKish arithmetization,
    Inner Product Argument on Pallas/Vesta curves, transparent setup.
  \item \textbf{plonky2} (Goldilocks): PLONKish arithmetization
    with FRI commitments over 64-bit Goldilocks field, transparent
    setup.
\end{itemize}

\subsection{Threat Model}

The proof system migration must preserve invariant I-07 (Proof
Soundness) from the Basis Network threat model (ZK-02): no
computationally bounded adversary can produce a valid proof for
an invalid state transition. During the dual-verification
transition period, the system must additionally guarantee that no
batch goes unverified---i.e., every submitted batch is validated
by at least one proof system at all times.
