\section{Conclusion}
\label{sec:conclusion}

This paper evaluated three recursive proof aggregation strategies for
compressing $N$ independent enterprise halo2-KZG validity proofs into a
single L1-verifiable proof on the Basis Network settlement layer.

\subsection{Key Findings}

\begin{enumerate}
  \item \textbf{Folding + Groth16 decider} achieves the strongest
    results across all metrics: $15.3\times$ gas reduction at $N{=}8$
    (from 3.36M to 220K gas), constant 128-byte proof size, sub-linear
    aggregation time (11.75\,s), and constant 1.6\,GB memory. The
    ${\sim}2N$-fold savings exceed the initial $N$-fold hypothesis
    because the Groth16 decider (220K gas) is cheaper than the original
    halo2-KZG verification (420K gas).

  \item \textbf{Binary tree accumulation} delivers exactly $N$-fold gas
    savings ($8.0\times$ at $N{=}8$) with production-proven tooling
    (Scroll, Axiom) but incurs 180\,s aggregation time due to sequential
    halo2-in-halo2 verification at each tree level.

  \item \textbf{SnarkPack} provides $7.5\times$ savings with the fastest
    aggregation (200\,ms at $N{=}8$) but requires Groth16-compatible
    inputs and incurs higher gas than the folding approach.

  \item Four aggregation invariants---soundness, independence
    preservation, order independence, and gas monotonicity---are
    identified as necessary conditions for production deployment.
\end{enumerate}

\subsection{Recommendations}

For the Basis Network zkEVM L2:

\begin{itemize}
  \item Deploy binary tree accumulation (Phase~1) for immediate
    $8\times$ gas savings using production-proven halo2 tooling.
  \item Transition to ProtoGalaxy folding + Groth16 decider (Phase~2)
    once the Sonobe library is audited, achieving $15.3\times$ savings
    with superior scaling properties.
  \item Formalize INV-AGG-1 through INV-AGG-4 in TLA+ (downstream
    Logicist task, checklist item~38) to ensure aggregation correctness
    is model-checked before implementation.
\end{itemize}

\subsection{Future Work}

\begin{itemize}
  \item \textbf{Stage 2 (Baseline):} Extend benchmarks to larger
    circuits (batch\_size 64, 256) with full stochastic analysis.
  \item \textbf{Stage 3 (Adversarial):} Test invalid proof injection,
    duplicate submissions, and wrong-circuit proofs to validate
    INV-AGG-1 and INV-AGG-2 under attack conditions.
  \item \textbf{Stage 4 (Ablation):} Compare Nova R1CS vs.\ ProtoGalaxy
    PLONKish folding, vary tree arity (2, 4, 8), and measure the gas
    delta of removing the Groth16 decider in favor of direct halo2-KZG
    final verification.
  \item \textbf{Cross-enterprise privacy:} Investigate selective
    disclosure mechanisms that allow auditors to verify individual
    enterprise proofs within an aggregated proof without revealing other
    enterprises' data.
\end{itemize}
