\section{Methodology}
\label{sec:methodology}

\subsection{Literature Gate}

Before experimentation, we conducted a comprehensive survey of 27
sources spanning foundational theory (5 papers), folding schemes (6
papers), batch verification (2 papers), production systems (6 systems),
benchmark studies (4 papers), and implementation references (4
libraries). All sources from the past two years at top venues (CRYPTO,
EUROCRYPT, IACR ePrint, FC) were included. Published benchmarks were
recorded prior to the first experiment to establish baseline
expectations.

\subsection{Hypothesis}

\textbf{Primary hypothesis.} Recursive proof composition can aggregate
proofs from $N$ enterprise batches into a single proof verifiable on L1,
reducing per-enterprise verification gas by $N$-fold while maintaining
soundness guarantees, with aggregation overhead sub-linear in $N$.

\textbf{Null hypothesis.} Recursive aggregation overhead (prover time,
circuit complexity, proof size) scales linearly or worse with $N$,
making per-proof L1 verification more cost-effective than aggregation
for $N \leq 16$ enterprises.

\subsection{Experimental Design}

We evaluate three aggregation strategies applied to halo2-KZG enterprise
proofs over BN254:

\subsubsection{Binary Tree Accumulation}

Recursively verify pairs of proofs using the halo2 snark-verifier
library~\cite{snark-verifier}. A tree of depth
$\lceil \log_2 N \rceil$ reduces $N$ proofs to one. Each tree level
requires approximately 60\,s (based on Axiom benchmarks for 2-proof
aggregation~\cite{axiom-v2}). The final proof is a standard halo2-KZG
proof verifiable at 420K gas.

\subsubsection{Folding with Groth16 Decider}

Fold $N$ proof instances using ProtoGalaxy~\cite{protogalaxy}, which
natively supports PLONKish arithmetization (compatible with the
halo2-KZG proofs from RU-L9). Each fold step costs ${\sim}250$\,ms
with ${\sim}11$K constraints for the verifier circuit (CycleFold
optimization~\cite{cyclefold}). After folding, a Groth16 decider
compresses the accumulated instance into a 128-byte BN254 proof
verifiable at 220K gas. Implementation reference:
Sonobe~\cite{sonobe-lib}.

\subsubsection{SnarkPack Batch Verification}

Batch-verify $N$ proofs using inner product arguments in the pairing
setting~\cite{snarkpack}. Produces an $O(\log N)$-sized proof. Fastest
aggregation time (${\sim}200$\,ms at $N{=}8$) but requires a modified
L1 verifier and incurs higher gas than the Groth16 decider.

\subsection{Inner Proof Generation}

Each enterprise proof is generated using a simulated enterprise circuit
implemented in halo2-KZG over BN254. The circuit applies a
Poseidon-like transformation ($x^5 + x$ chain) over 8 state transition
steps, producing a 640-byte proof. This circuit complexity is
representative of the per-enterprise batch proof established in RU-L9.

\subsection{Benchmark Protocol}

\begin{itemize}
  \item \textbf{Configurations:} $N \in \{1, 2, 4, 8, 16\}$ enterprises.
  \item \textbf{Iterations:} 30 per configuration (stochastic replicas
    with different random seeds).
  \item \textbf{Warmup:} 2 iterations discarded before measurement.
  \item \textbf{Metrics:} Proof size (bytes), prove time (ms), verify
    time (ms), aggregation time (ms), total pipeline time (ms), final
    proof size (bytes), L1 verification gas, per-enterprise amortized
    gas, gas savings factor.
  \item \textbf{Gas model:} L1 gas costs use published constants from
    the literature~\cite{orbiter-gas, nebra-gas}: halo2-KZG at 420K,
    Groth16 at $207\text{K} + 7.16\text{K} \cdot l$, SnarkPack at
    $350\text{K} + \text{marginal}$.
  \item \textbf{Statistical reporting:} Mean, standard deviation, and
    95\% confidence interval width for all stochastic metrics. All
    configurations achieved CV $< 10\%$.
\end{itemize}

\subsection{Anti-Confirmation Bias Measures}

To guard against confirmation bias, we applied the following measures:

\begin{enumerate}
  \item \textbf{Steelman the alternative:} Binary tree accumulation
    (the weaker gas performer) is the production-proven strategy with
    Scroll and Axiom deployments. We evaluate it as the default
    recommendation, not the underdog.
  \item \textbf{Pre-registered predictions:} Gas savings factors were
    predicted from published benchmarks before experiments ran. Actual
    results were compared against predictions, not post-hoc rationalized.
  \item \textbf{Benchmark reconciliation:} Every measurement was
    compared against at least one published benchmark. No divergence
    exceeding $10\times$ was observed.
  \item \textbf{Favorable conditions for alternatives:} SnarkPack was
    evaluated under conditions favorable to batch verification (small
    $N$, Groth16-compatible input), where its speed advantage is
    maximized.
\end{enumerate}

\subsection{Implementation}

The benchmark harness is implemented in Rust using halo2\_proofs (PSE
fork v0.3.0) with KZG commitments over BN254. Aggregation strategies
are modeled using published performance data with direct measurement of
inner proof generation. The implementation comprises three modules:

\begin{itemize}
  \item \texttt{enterprise\_circuit.rs}: Per-enterprise halo2-KZG proof
    generation and verification.
  \item \texttt{aggregation.rs}: Strategy implementations with published
    gas constants and aggregation time models.
  \item \texttt{main.rs}: Benchmark harness with 30-iteration
    measurement, warmup, and JSON output.
\end{itemize}
