\section{Related Work}
\label{sec:related}

We organize related work into four categories: foundational recursive
SNARK theory, folding schemes, batch verification and aggregation, and
production systems.

\subsection{Recursive SNARK Composition}

Bitansky et al.~\cite{bitansky-pcd} introduced proof-carrying data
(PCD), establishing the theoretical foundation for recursive SNARK
composition. Ben-Sasson et al.~\cite{ben-sasson-cycle} demonstrated
practical recursion using cycles of elliptic curves (MNT4/MNT6),
enabling native-field verification. Bowe, Grigg, and
Hopwood~\cite{halo} introduced accumulation schemes that defer
polynomial commitment checks rather than verifying them in-circuit,
eliminating the need for trusted setup in the recursive step.
B\"{u}nz et al.~\cite{bunz-accumulation} formalized the accumulation
framework, generalizing the Halo approach to a broader class of
proof systems. The inner product argument techniques of
TIPP/MIPP~\cite{bunz-mipp} provided the algebraic foundation for
SnarkPack's batch verification.

\subsection{Folding Schemes}

Nova~\cite{nova} introduced folding for relaxed R1CS, achieving
incrementally verifiable computation (IVC) with a verifier circuit of
approximately 10K constraints and constant memory regardless of the
number of folding steps. CycleFold~\cite{cyclefold} reduced the
secondary curve circuit from ${\sim}10$K to ${\sim}1.5$K constraints.
SuperNova~\cite{supernova} extended folding to non-uniform IVC,
supporting per-opcode circuits without a universal circuit tax.
ProtoStar~\cite{protostar} and ProtoGalaxy~\cite{protogalaxy} adapted
folding to PLONKish arithmetization, enabling native support for custom
gates and lookup tables. HyperNova~\cite{hypernova} unified R1CS,
PLONKish, and AIR under the Customizable Constraint System (CCS)
abstraction, achieving multi-folding with a single multi-scalar
multiplication per fold step.

\subsection{Batch Verification and Aggregation}

SnarkPack~\cite{snarkpack} batches $N$ Groth16 proofs into a single
$O(\log N)$-sized proof using inner product arguments in the pairing
setting, achieving verification of 8,192 proofs in 8.7\,s with proof
size under 1.8\,KB. SnarkFold~\cite{snarkfold} applies Nova-style
folding to SNARK verification circuits, reporting 4.5\,ms verification
for 4,096 aggregated proofs.

\subsection{Production Systems}

Six production systems provide reference benchmarks for aggregation
performance:

\begin{itemize}
  \item \textbf{Polygon zkEVM}~\cite{polygon-zkevm} uses eSTARK to
    FFLONK binary tree aggregation, achieving ${\sim}350$K gas for the
    final verification.
  \item \textbf{zkSync Era (Boojum)}~\cite{zksync-era} employs a custom
    PLONK+FRI system with 15 recursive circuits and a PLONK-KZG wrapper,
    verifying at ${\sim}460$K gas.
  \item \textbf{Scroll (Darwin)}~\cite{scroll-darwin} implements a
    chunk-batch-bundle hierarchy using halo2-KZG, achieving 34\% gas
    reduction via bundling.
  \item \textbf{Nebra UPA}~\cite{nebra-upa} provides universal proof
    aggregation, reaching ${\sim}18$K gas per proof at $N{=}32$.
  \item \textbf{Axiom V2}~\cite{axiom-v2} uses halo2-KZG with the
    snark-verifier library for recursive aggregation at 420K fixed gas
    on Ethereum mainnet.
  \item \textbf{Taiko (Raiko)}~\cite{taiko-raiko} implements a
    multi-proof hybrid (SGX+ZK) but is not primarily focused on proof
    aggregation.
\end{itemize}

Our work differs from these systems in its enterprise-specific focus:
rather than aggregating batches within a single rollup, we aggregate
independent proofs from $N$ separate enterprise chains, each with its
own circuit and state. This introduces requirements for independence
preservation and order independence that do not arise in single-rollup
aggregation.
